\documentclass[../main.tex]{subfiles}
\graphicspath{{\subfix{../images/}}}

\begin{document}

\hypertarget{introduction}{%
\chapter{Introduction}\label{introduction}}

Although computers had already emerged in their analog form during the
19th century, their significant proliferation began in the mid-20th
century. 1941 marked the creation of Z3, the initial operational digital
computer developed by Konrad Zuse. The distinctive feature of this
machinery lies in its programmability through the utilization of punch
cards. The source code, which comprises instructions that can be
comprehended and executed by the computer, has the potential to be
stored in an analog form.

After eighty years, these computing machines have become ubiquitous in
our surroundings. Despite the rise in their usefulness, the fundamental
structure remains unchanged: a collection of software programs dictate
the behavior of physical circuits or hardware. As a result of the rise
in utility costs, there was a corresponding increase in the number of
users. Initially, a part of the users exhibit benign behavior, utilizing
technology for its intended purpose. Conversely, there exist malicious
actors who aim to acquire either material or immaterial (such as
reputation or recognition) advantages, through the exploitation of
technology, including its software sections.

The susceptibility of software to attacks can be attributed to the fact
that it is developed by individuals who are inherently prone to errors.
The liability for the error should not rest solely on the individual who
authored the code, but rather, measures for validation could be
implemented to safeguard the written code. The techniques employed
comprise either a manual or automated procedure. One instance of the
former is a security code review, wherein a proficient engineer assesses
the security posture of a codebase. Regarding the latter, it is possible
to incorporate security linters, scanners, and fuzzers into integrated
development environments (IDEs), continuous integration/continuous
deployment (CI/CD) pipelines, or staging environments.

The Cyber Grand Challenge was convened in 2016 by the Defense Advanced
Research Projects Agency (DARPA), an agency dedicated to research and
development under the purview of the United States Department of
Defense. The event involved a competition among automated systems, which
engaged in a demonstration of their accuracy and effectiveness by
engaging in warfare with one another. Termed cyber reasoning system (CRS),
their primary objective is to identify weaknesses in the binaries that
are executed by both them and their opponents simultaneously.
Participants in the Capture The Flag (CTF) competition had the opportunity to
accumulate points through the automated generation of exploits that
could be executed against their opponents, as well as by patching their
binaries to safeguard against potential attacks from adversaries.

Currently, the three leading contenders are deemed to be
state-of-the-art. Mayhem \cite{mayhem}, the winner, was turned into a commercial
solution. Although Shellphish \cite{shellphish} was released as an open-source
project, the open-source environment has not undergone any
significant changes. This matter serves as the initial point for two
theses, specifically the present one and another \cite{ghenea} authored by Claudiu
Ghenea. The authors present OpenCRS, an open-source system for
cyber reasoning that is capable of identifying, exploiting, and
remedying vulnerabilities in C-based executables designed for
\texttt{i386} and in Executable and Linkable Format (ELF).

\hypertarget{objectives}{%
\section{Objectives}\label{objectives}}

The primary aim of this manuscript is to explain and assess four
components of the cyber reasoning system that were developed by the
author:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  The dataset module, responsible for managing test suites of vulnerable
  programs;
\item
  The attack surface approximation module, which identifies the
  executable's interaction with the environment;
\item
  The vulnerability detection module, which generates input that
  triggers a vulnerability in the executable; and
\item
  The automatic exploit generation module, which produces a functional
  exploit for each identified vulnerability.
\end{enumerate}

In addition, the document will briefly outline how OpenCRS works, an
aspect that will be handled in greater depth by the other thesis that we
mentioned.

\hypertarget{structure}{%
\section{Structure}\label{structure}}

The thesis is divided into eight chapters.

\begin{itemize}
    \item Chapter \ref{introduction}: This one described the present context for software security, cyber reasoning systems, and the objective of this publication.
    \item Chapter \ref{opencrs}: The subsequent chapter will expound upon OpenCRS, providing a comprehensive overview of its architecture and the limitations that have been established for this particular system.
    \item Chapters \ref{dataset-module} to \ref{automatic-exploit-generation-module}: The next four chapters will expound on each module that was referenced in the enumerated list aforementioned. Each module provides a detailed description of the theoretical aspects of binary analysis subfields that are implemented, including the state-of-the-art in each area. Additionally, the architecture and implementation of each module, as well as the technologies employed, are thoroughly discussed.
    \item Chapter \ref{testing-and-assesment}: Chapter seven of the module focuses on testing and evaluation, providing precise data that demonstrates the accuracy of the results.
    \item Chapter \ref{conclusion_future}: The final chapter provides a summary of the thesis content and draws conclusions based on the findings presented. In addition, it offers recommendations for future enhancements and modifications to the OpenCRS platform and its associated modules.
\end{itemize}

\end{document}